Maximum Entropy Prediction Markets

### 2014-02-06 WriteUp/maxentpredmkts.tex
minor edits
Revert "Progress on information markets section"
This reverts commit 4b4cb0559fce0de653a6299aa65828367c5d65e4.
first draft of intro

\section{Introduction}
%------------------------------------------------------------
%Outline
%------------------------------------------------------------
%-introduction [Sindhu; Thursday]
%-literature review
%- maxent scoring rule [Sebastien]
%	- generalized lmsr
%	- exp family review
%	[- does modularity (a la Hanson) still hold?]
%- prediction market setup [Sebastien]
%	- MLE with risk neutrality
% 	need a definition of term 'exponential family prediction market'
%- conjugate prior market [Sindhu; Wednesday]
%	- also on hyperparam?
%- exp-util [Sebastien]
%- equilibrium for repeated trades [Sindhu; Tuesday] done
% 	[- rate of convergence?]
%- budgets and adversarial markets [Sindhu; Monday] done
% [any liquidity results?] [Sindhu: weekend]
% conclusion
% bibliography

@@

We have dual goals in this paper. On the one hand, we highlight a structural similarity between prediction markets and exponential families. We see this as the syntax of our prediction market mechanism. On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments. For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
%\section{Notation and Definitions}
%convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.

>>>>

\section{Introduction}
%------------------------------------------------------------
%Outline
%------------------------------------------------------------
%-introduction [Sindhu; Thursday]
%-literature review
%- maxent scoring rule [Sebastien]
%	- generalized lmsr
%	- exp family review
%	[- does modularity (a la Hanson) still hold?]
%- prediction market setup [Sebastien]
%	- MLE with risk neutrality
% 	need a definition of term 'exponential family prediction market'
%- conjugate prior market [Sindhu; Wednesday] done
%	- also on hyperparam?
%- exp-util [Sebastien]
%- equilibrium for repeated trades [Sindhu; Tuesday] done
% 	[- rate of convergence?]
%- budgets and adversarial markets [Sindhu; Monday] done
% [any liquidity results?] [Sindhu: weekend]
% conclusion
% bibliography

@@

%- prediction market construction
%	- exponential failies
%	- scoring rules
%	- information market
%- Traders
%	- aggregation process
%		- bayesian traders
%		- frequentist
%	- utility
%		- linear
%		- exponential
%	- budgets
%- multiple trades
%- multiple traders
Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event. Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event. The traders are assumed to be strategic in the sense that every trader wants to maximize his profit by participating in the market. Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.

Traders are allowed to report their beliefs on the outcome of the event by allowing them to buy and sell securities whose value depends on the outcome of this future event. This will effect the state of the market, thus updating the predictive probabilities for the event. For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins. Traders can see the past history of trades, so the price at which a current trader is willing to buy and sell these securities can be interpreted as an aggregate ``market probability'' of a particular candidate winning the election.

Currently, most deployed prediction markets have binary or finite discrete outcome spaces. This means that the outcome of every event can be characterized by a finite set of mutually exclusive exhaustive outcomes. The securities in these market are usually tied to each outcome so that buying (respectively, selling) a security related to an outcome, indicates a belief in an increased (respectively, decreased) probability of occurrence of the outcome.

Directly applying techniques used for binary outcome spaces to large outcome spaces is computationally infeasible. In fact, it has been shown that computing and updating prices for combinatorial outcomes spaces is $\#$P-hard \cite{Chen08}. Infinite outcome spaces pose a particular challenge in prediction markets since it is not clear how to define securities in such spaces. The most direct way of discretizing the space may lead to loss of information from the traders. Extending prediction market techniques to this space has been a subject of recent interest. Various techniques to bound market maker loss in both a combinatorial and infinite outcome space have been studied (see, for instance, \cite{Abernethy13,OthmanS11}).

One popular form of prediction markets is the market scoring rule~\cite{hanson03}. A market scoring rule considers all trades as a single chronological sequence. Traders earn rewards proportional to the incremental reduction in prediction loss caused by their trades in comparison to the previous trade. In other words, their rewards depend on the change in market probabilities caused by their trade, as well as on the eventual outcome. Thus, each trader has an 
incentive to minimize the prediction loss. In this format, the {\em market maker} who runs the market can suffer an overall loss, but Hanson~\cite{hanson03} showed that, for market scoring rules on finite outcome spaces, the loss of the market maker can be bounded.

In this paper, we propose a general version of the popular logarithmic scoring rule extended to continuous outcome spaces. Our version of the scoring rule uses parametric distributions to elicit traders' expectation on some function (called statistics) of the data. It turns out that the maximum entropy distribution given the statistics are  exponential family distributions.

We have dual goals in this paper. On the one hand, we highlight a structural similarity between prediction markets and exponential families. We see this as the syntax of our prediction market mechanism. On the other, this formulation has rich semantics as well: it allows for analysis of the evolution of the market under various  models of trader behavior.

Traders with limited budgets are constrained in how much they influence the market. Their purchases may not reflect their true beliefs. We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.

We also analyze the behavior of traders with prior exposure in the market. Our results formalize the intuitive notion that private belief and financial exposure are just two sides of the same coin; trades in a prediction market can be seen as updating privately held beliefs.

For risk averse traders, we show how a single trader interacts with the market as well as when multiple traders achieve equilibrium. The eventual market state turns out to be a weighted combination of trader beliefs and initial market state; the weights depend on the risk aversion parameter of the individual traders. Trader behavior varies depending on whether they assimilate information as Bayesians or as frequentists.  Interestingly, we show that although these two kinds of traders pick the eventual market state as a convex combination of their private belief and current market state, they do so in dual spaces.

%\section{Notation and Definitions}
%convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
